---
author: "Elizabeth Cruz"
date: "18 de Fevereiro de 2017"
output: html_document
---

<center><h1>What is ggplot2?</h1></center>
 
<center><h2>A Data Visualization package for R</h2></center>
<br>

- An implementation of **Grammar of Graphics** - a general scheme for data visualization which breaks up graphs into semantic components such as layers (think "noun", "verb", "adjective" for graphics)

- The basic concept is that you combine **plot building blocks** and combine them to create just about any kind of graphical display you want.

<br>
<center><h4>Data + Mapping + Geometric Object + <small>Stat + Coordinate Systems + Position Adjustments + Faceting + etc.</small></h4></center>

<br>
<h5>**Example:** ggplot(dataframe, aes(x=, y=)) + geom_point() + facet_grid()</h5>

***
                             
###Data Frame###
- Refers to "tabular" data
- List of variables (columns) of the same number of rows with unique row names


###Aesthetics Mappings                                    
* Determine how data is mapping                        
* Means "something you can see"                        
* Examples include:                                         
    + Position (on the x and y axes)                        
    + Color/fill                                            
    + Shape (of points)                     
    + Linetype                        
    + Size                   

###Geoms
* What you see in the plot
* Examples include:
    + Points
    + Lines
    + Bars

```{r, echo = FALSE}
knitr::include_graphics("https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf")
```

<br>

####There is qplot and ggplot
- Think of qplot() as a Quick Plot.
- Think of ggplot() as a muche more Complex Plot

### How to load ggplot2 and dataset?
```{r}
#some scaling
if("repr" %in% rownames(installed.packages()) == FALSE) {install.packages("repr")}
library(repr)
options(repr.plot.width=4, repr.plot.height=4)

#loading ggplot2 package
library(ggplot2)
```
``` {r}
#lets choose the dataset we are going to work on
library(MASS)

##Structure of the dataset UScereal
str(UScereal)
```
You can always type: ?help(USCereal)
```{r}
#some lines
head(UScereal)

options(repr.plot.width=4, repr.plot.height = 2.91)

#Basic Histogram without ggplot2
hist(UScereal$protein)
```
<br>

### Basic Plot
```{r}
options(repr.plot.width = 4, repr.plot.height = 4)

plot(UScereal$sugars, UScereal$calories)
title("Sugars and calories in US cereals")
points(sugars ~ calories, col="red", data=UScereal)
```
<br>

**EXERCÍCIO 1:** À semelhança do histograma acima, construir um histograma que apresente a contagem de proteínas.
<br>

***

## The ggplot
```{r}
g <- ggplot(UScereal, aes(sugars, calories))
```
This means that x is "sugars" and y is "calories".

"Sugars" and "Calories" are column names of the Data Frame UScereal.

<br>

### SCATTER PLOTS - Let's talk about them!!
```{r}
g + geom_point()
```

Let's practice modifying aesthetics now.

```{r}
g + geom_point(aes(color = I("brown")), size=5, alpha = I(1/5))
```

To make a semi-transparent colour you can use the alpha aesthetic, which takes a value between 0 (completely transparent) and 1 (complete opaque).

It's often useful to specify the transparency as a fraction, e.g., 1/10 or 1/20, as the denominator specifies the number of points that must overplot to get a completely opaque colour.

We can see that some circles are more transparent than others, because they correspond to a higher frquency.

```{r}
g + geom_point(size=5, alpha = I(1/3), aes(color = vitamins)) + 
    labs(title = "Working with Cereals!", x = "Sugars per portion", y= "Calories per portion")
```

**EXERCÍCIO 2:**

- Utilizar o dataset "airquality" que vem com o R
- Fazer um scatter plot com Day vs Ozone
- Deixar a cor ser Wind
- Introduzir alguma transparência
- Definir um tamanho para os pontos
- Dar um título ao gráfico e dar nomes aos eixos

***

```{r}
g <- ggplot(airquality, aes(Day,Ozone))
```
```{r, warning=FALSE}
g + geom_point(size=4, alpha=I(1/3), aes(color=Wind)) + labs(title="Day vs. Ozone", ylab="Ozono", xlab="Dia da semana")
```
<br>

###geom_smooth
geom_smooth fits a curve between two variables.

*Method LOESS:* LOcally WEighted Scatter-plot Smoother (known as locally weighted polynomial regression) or Method LM: Linear Model

```{r}
g <- ggplot(UScereal, aes(sugars, calories))
g + geom_point() + geom_smooth(method="loess")
```
```{r}
g + geom_point() + geom_smooth(method = "lm")
```
```{r}
g + geom_point(size=5, alpha = 1/2, aes(color = vitamins)) + 
    labs(title = "Working with Cereals Again!", x = "Sugars per portion", y= "Calories per portion") + 
    geom_smooth(size=4, linetype=1, method="lm", se=FALSE) + theme_bw()
```

**EXERCÍCIO 3:** 

- Experimentar construir o gráfico acima mas brincando com o tipo de linha e tamanho. 
- Tentar que a linha seja a tracejado e mais fina.

```{r}

```


***

###Boxplots
```{r}
g <- ggplot(UScereal, aes(x=mfr, y=calories))
### G=General Mills, K=Kelloggs, N=Nabisco, P=Post, Q=Quaker Oats, R=Ralston Purina.
```
```{r}
g + geom_boxplot(aes(fill=mfr, color=I("black"))) + 
    ylim(75, 240) +
    labs(title = "Working with Cereals Again!", x = "Manufacturer", y= "Calories per portion") + 
    theme_bw()
### mfr=manufacturer
```

In a boxplot, we can see the change in the median and adjacent quartiles.
<br>

###Histogram and Density plots
Histogram and density plots show the distribution os a single variable.

The following code shows the distribution of calories:
```{r}
g <- ggplot(UScereal, aes(calories))
g + geom_histogram(aes(y=..density.., fill=I("blue"))) + 
    geom_density(aes(color=I("brown"))) + 
    labs(title="Contagem de Calorias", x="Calorias", y="Contagem") + 
    theme_bw()
```

###Waht is the Bin Width?
```{r}
g <- ggplot(UScereal, aes(calories))
g + geom_histogram(aes(fill=I("blue")))
```

We get a message about the binwidth defaulting to 30.

When you make a histogram, you're essentially making a bar chart that shows how many data points fit within a certain range. That range is the bin width.

So in a histogram, data is grouped into "bins".

To construct a histogram, the first step is to "bin" the range of values—that is, divide the entire range of values into a series of intervals—and then count how many values fall into each interval. The bins are usually specified as consecutive, non-overlapping intervals of a variable.

Lets try to calculate the bin width.

We simply find the range (maximum - minimum) and divide that by how many bins we'd like to have.

A rough guideline for the number of bins is: take the square root of the number of data values you have.

```{r}
range_calories = max(UScereal$calories) - min(UScereal$calories)
range_calories
```
```{r}
number_bins = sqrt(nrow(UScereal))
number_bins
```
```{r}
g + geom_histogram(binwidth = 48, aes(fill=I("blue")))
```

No more messages in red!

```{r}
g + geom_histogram(binwidth = 48, aes(color=mfr, fill=mfr)) + 
    labs(title="Calories Count", x="Calories", y="Frequency") + 
    theme_bw()
```

Too complicated! Lets chack a boxplot of it.

```{r}
ggplot(UScereal, aes(x=mfr, y=calories)) + 
    geom_boxplot(aes(fill=mfr, color=I("black"))) + 
    ylim(75, 240) +
    labs(title = "Working with Cereals Again!", x = "Manufacturer", y= "Calories per portion") + 
    theme_bw()
```
<br>

###Time series with line and path plots
Shows how a single variable change over time.

Because there is no time variable in the USCereal data, we use the economics dataset wich describes unemployment in a certain population over time.

```{r}
str(economics)
```
```{r}
g <- ggplot(economics, aes(date, unemploy))
g + geom_line(aes(color=I("blue"))) + 
    labs(title="Our First Time Series", x="Date", y="Unemployment in thousands") +
    theme_bw()
```

```{r}
options(repr.plot.width=4, repr.plot.height=2.9)

g <- ggplot(economics, aes(date, unemploy/pop))
g + geom_line(aes(color=I("blue"))) + 
    labs(title="Our Second Time Series", x="Date", y="Unemployment/Population") + theme_bw()
```

Some other themes: Balck and White, Grey Theme and The FiveThirtyEight Theme

```{r}
library(ggthemes)

options(repr.plot.width=4, repr.plot.height=2.9)

g <- ggplot(economics, aes(date, unemploy/pop))

g + geom_line(aes(color=I("blue"))) + 
    labs(title="Our Second Time Series", x="Date", y="Unemployment/Population") + theme_bw()

g + geom_line(aes(color=I("blue"))) + 
    labs(title="Our Second Time Series", x="Date", y="Unemployment/Population") + theme_grey()

g + geom_line(aes(color=I("blue"))) + 
    labs(title="Our Second Time Series", x="Date", y="Unemployment/Population") + theme_fivethirtyeight()
```

###Faceting
We have already discussed using aesthetics (colour and shape) to compare subgroups, drawing all groups on the same plot.

Faceting takes an alternative approach: It creates tables of graphics by splitting the data into subsets and displaying the same graph for each subset in an arrangement that facilitates comparison.

The default faceting method in GGplot() creates plots arranged on a grid specified by a faceting formula which looks like row_var ~ col_var.

You can specify as many row and column variables as you like, keeping in mind that using more than two variables will often produce a plot so large that it is difficult to see on screen. To facet on only one of columns or rows, use . as a place holder.

For example, row_var ~ . will create a single column with multiple rows.

For faceting we will use a different dataset called Diamonds.

```{r}
str(diamonds)
```

```{r}
head(diamonds)
```

The following code illustrates this technique with a histogram showing the distribution of carats conditional on color.

```{r}
options(repr.plot.width=5, repr.plot.height=6)

ggplot(diamonds, aes(carat)) + 
    geom_histogram(binwidth = 0.1, aes(fill=I("blue"))) + 
    facet_grid(color ~.) + 
    coord_cartesian(xlim = c(0, 3), ylim = c(0, 1800)) + 
    theme_bw() + 
    labs(x="Carats", y="Frequency")
```

```{r}
diamonds_small <- diamonds[sample(1:nrow(diamonds), 200, replace=FALSE),]

str(diamonds_small)
```

Now, lets try it with some scatter plots.

```{r}
options(repr.plot.width=4, repr.plot.height=4)

g <- ggplot(diamonds_small, aes(x=carat, y=price, color=color)) + 
    geom_point(shape=1) +
    labs(title="Smaller Scatter Plot", x="Carat", y="Price") + 
    theme_fivethirtyeight()
g
```

```{r}
options(repr.plot.width=5, repr.plot.height=6)

# Divide by levels of "color", in the vertical direction
g <- ggplot(diamonds_small, aes(x=carat, y=price, color=color)) + 
    geom_point(shape=1) +
    labs(title="Smaller Scatter Plot", x="Carat", y="Price") + 
    theme_bw() +
    facet_grid(color ~ .)
g
```

```{r}
options(repr.plot.width=7, repr.plot.height=6)

# Divide by levels of "color", in the horizontal direction
g <- ggplot(diamonds_small, aes(x=carat, y=price, color=color)) + 
    geom_point(shape=1) +
    labs(title="Smaller Scatter Plot", x="Carat", y="Price") + 
    theme_bw() +
    facet_grid(. ~ color)
g
```

```{r}
options(repr.plot.width=7, repr.plot.height=6)

# Divide with "cut" vertical, "shape" horizontal
g <- ggplot(diamonds_small, aes(x=carat, y=price, color=color)) + 
    geom_point(shape=1) +
    labs(title="Smaller Scatter Plot", x="Carat", y="Price") + 
    theme_grey() +
    facet_grid(cut ~ color)
g
```

##Plotly - Interactive ggplot
In RStudio.

##Sumário para recapitular:
- ggplot2 baseado na "Gramática de Gráficos"

- str() e head() para inspeccionar base de dados

- os 2 atributos do ggplot2 mais importantes são: Aesthetics e Geoms

- geom_point: Gráfico de Pontos
- geom_smooth: Curva
- geom_boxplot: Gráfico de "Caixas"
- geom_hist e geom_density: Histograma e Densidade
- geom_line: Time Series

- facet_grid: "multigráfico"

- theme_bw/theme_grey/theme_fivethirtyeight

- ggplotly: pacote Plotly para interactividade